---
title: "p8105_hw3_sc5352"
author: "Chrissy Chen"
date: "2023-10-14"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 2

```{r}
library(p8105.datasets)
library(tidyverse)
library(dplyr)
library(ggplot2)
data("brfss_smart2010")
brfss_smart = 
  brfss_smart2010 |>
  janitor::clean_names() |>
  rename(state = locationabbr, state_location = locationdesc) |>
  filter(topic == "Overall Health",
         response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) |>
  mutate(
    response = factor(response, 
         levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
         ordered = TRUE))
brfss_smart
```
##### In 2002, which states were observed at 7 or more locations? What about in 2010?
```{r}
brfss_smart_2002 =
  brfss_smart |>
  filter(year == 2002) |>
  distinct(state,state_location)|>
  count(state) |>
  filter(n >= 7) 
brfss_smart_2002

brfss_smart_2010 =
  brfss_smart |>
  filter(year == 2010) |>
  distinct(state,state_location)|>
  count(state) |>
  filter(n >= 7)
brfss_smart_2010
```

2002: CT, FL, MA, NC, NJ, PA  
2010: CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA  

##### Make a plot showing a line for each state across years

```{r}
brfss_smart_Excellent = 
  brfss_smart |>
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarize(mean_data_value = mean(data_value, na.rm = TRUE)) |>
  ggplot(aes(x = year, y = mean_data_value, color = state)) + 
    geom_point() + geom_line() +
  labs(
    title = "Year vs Mean Data Value",
    x = "Year",
    y = "Mean Data Value"
    )
brfss_smart_Excellent
    
```
There is no clear trend for the plot. The variation is large. However, the mean values for these states focus on 20-25.     

##### Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_smart_NY = 
  brfss_smart |>
  filter((year == 2006|year == 2010), state=="NY") 
brfss_smart_NY

ggplot(brfss_smart_NY, aes(x = response, y = data_value)) + 
  geom_boxplot() + 
  facet_grid(.~year) +
  labs(
    title = "Response vs Data Value",
    x = "Response",
    y = "Data Value"
  )

brfss_smart_NY
```


### Problem 3

```{r}
covar = 
  read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  drop_na() |>
  mutate(sex = recode(sex, 
                      `1` = "male", 
                      `2` = "female"),
         education = recode(education, 
                            `1` = "Less than high school", 
                            `2` = "High school equivalent", 
                            `3` = "More than high school"))

accel = 
  read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  drop_na() |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "minute_value"
  )

combine_data = 
  left_join(covar, accel, by = "seqn") |>
  filter(age >= 21) |>
  #drop_na() |>
  mutate(education = factor(education,
                            levels = c("Less than high school", 
                                       "High school equivalent",
                                       "More than high school"),
                            ordered = TRUE))
```

##### Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items. 
```{r}
combine_data_readfriendly = 
  covar |>
  filter(age >= 21) |>
  group_by(sex, education) |>
  tally()|>
  pivot_wider(names_from = sex,values_from = n)|>
  knitr::kable()
combine_data_readfriendly
```
`Female` with `More than high school` has the highest number with 59. However, `Female` with `High school equivalent` has the lowest number with 23.

```{r}
combine_data_visual = 
  combine_data |>
  ggplot(aes(x = education, y = age, color = sex)) + 
  geom_boxplot() +
  labs(
    title = "Age vs Education",
    x = "Education",
    y = "Age"
  )
combine_data_visual
```
In "less than high school" group, female and male have similar ages. In "high shcool equivalent" group, the age divergence is the largest since female are older than the male on avergae. In "More than high school" group, the male are slightly older than the female.  

##### Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}
combine_data_new = 
  combine_data |>
  group_by(seqn) |>
  mutate(total_act = sum(minute_value)) |>
  filter(minute == "min1") |>
  select(-minute, -minute_value)

ggplot(combine_data_new, aes(x = age, y = total_act, color = sex)) + 
  geom_point() + 
  facet_grid(.~education) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Age vs Total Activity",
    x = "Age",
    y = "Total Acitivity"
  )

```
In general, as age increases, both men and women decrease their total activities. However, in the `high school equivalent` group, women increase their acitvities before 40 first and then decrease. Both men and women with `less than high school` education have the highest total activities among three groups.

##### Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}
combine_data_plot =
  combine_data |>
  mutate(minute = as.numeric(str_replace_all(minute, "min",""))) |>
  ggplot(aes(x = minute, y = minute_value, color = sex)) + 
  geom_point(alpha = 0.5) + 
  facet_grid(.~education) +
  geom_smooth() +
  labs(
    title = "Minute vs Minute Acitivity",
    x = "Minute",
    y = "Minute Acitivity"
  )
combine_data_plot
```
In `More than high school` group, the range of `Minute Acitivity` is the largest, compared to other groups. The smooth line for three panels look similar.
